Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-25 paper/paper.tex
edits to section 6

\subsection{Probabilistic Forms}

@@

The remaining Anglican special forms support carrying values in
the hidden program state. \texttt{mem} implements stochastic
function  memoization (Section~\ref{sec:mem}). \texttt{store}
stores a value for a tag in the program state, which can be
later retrieved by the tag using \texttt{retrieve} during the same
program execution.

@@

Before delving into implementation details, we provide a case
study of using the Anglican language and environment. This case
study takes a problem for which a solution is not immediately
obvious, \textit{the Deli dilemma}, and guides through writing
an Anglican program for the problem, executing the program, and
post-processing results.

The program presented in this section is intentionally short and
simple. Anglican is capable of compiling and running elaborated
programs handling large amounts of data. Advanced examples of
Anglican programs and inference can be found in literature on
applications of probabilistic programming~\cite{PLW15, P16, MPT+16}.

@@

Imagine that we are facing the following dilemma:
\begin{quote}
A customer wearing round sunglasses came to a deli at 1:13pm,
and grabbed a sandwich and a coffee. Later on the same day, a
customer wearing round sunglasses came at 6:09pm and ordered a
dinner.  Was it the same customer?
\end{quote}
Additionally, we know that:
\begin{itemize}
    \item There is an adjacent office quarter, and it takes
        between 5 and 15 minutes on average to walk from an
        office to the deli, where different average times are  for
        different buildings in the quarter.
    \item Depending on traffic lights, the walking time varies by about 2
        minutes.
    \item The lunch break starts at 1:00pm, and the workday ends at 6:00pm.
    \item The waiter's odds that this is the same customer are 2 to 1.
\end{itemize}

@@

We want to formalize the dilemma in an Anglican query, based on
the knowledge we have. Let us formalize the knowledge in
Clojure (the times are in minutes). First, we encode our prior
information which holds true independently of the customer's
visit:

\begin{lstlisting}[style=default]
(def mean-time-to-arrive
     "average time to arrive"
     10.)
(def sd-time-to-arrive
     "standard deviation of arrival time"
     3.)
(def time-sd
     "walking time deviation"
     1.)
\end{lstlisting}

@@

Performing inference on query \texttt{deli} computes the
probability that the same customer visited the deli twice, as
well as probability distributions of average arrival times for
both cases.

>>>>

\subsection{Special Forms}

@@

The remaining Anglican special forms support carrying values in
the hidden program state. \texttt{mem} implements stochastic
function  memoization (Section~\ref{sec:mem}). \texttt{store}
stores a value for a tag in the program state, which can be
later retrieved using \texttt{retrieve} during the same
program execution.

@@

Before delving into implementation details, we discuss a case study
of the use of the Anglican language and environment. This case study
takes an inference problem for which a solution is not immediately
obvious, \textit{the Deli dilemma}, and discusses how this problem
can be solved by writing an Anglican program, executing the program,
and post-processing results.

The program presented in this section is intentionally short and
simple. Anglican is capable of compiling and running elaborate
programs and handling large amounts of data. Advanced examples of
Anglican programs and inference can be found in literature on
applications of probabilistic programming~\cite{PLW15, P16, MPT+16}.

@@

Imagine that we are facing the following dilemma:
\begin{quote}
A customer wearing round sunglasses came to a deli at 1:13pm,
and grabbed a sandwich and a coffee. Later on the same day, a
customer wearing round sunglasses came at 6:09pm and ordered a
dinner.  Was it the same customer?
\end{quote}
Additionally, we know that:
\begin{itemize}
    \item There is an adjacent office quarter, and it takes
        between 5 and 15 minutes on average to walk from an
        office to the deli, where different average times are  for
        different buildings in the quarter.
    \item Depending on traffic lights, the walking time varies by about 2
        minutes.
    \item The lunch break starts at 1:00pm, and the workday ends at 6:00pm.
    \item Based on the similarity of appearance, the waiter assesses the odds that this is the same customer as 2 to 1.
\end{itemize}

@@

We want to formalize the dilemma as an Anglican query, based on
the knowledge we have. Let us formalize the knowledge in
Clojure (the times are in minutes). First, we encode our prior
information which holds true independently of the customer's
visit:

\begin{lstlisting}[style=default]
(def mean-office-distance
     "average time to arrive"
     10.)
(def sd-time-to-arrive
     "standard deviation of arrival time"
     3.)
(def time-sd
     "walking time deviation"
     1.)
\end{lstlisting}

@@

Performing inference on the query \texttt{deli} computes the
probability that the same customer visited the deli twice, as
well as probability distributions of average arrival times for
both cases.
